A significance test for forward stepwise model selection

### 2014-05-15 paper/paper.tex
Fixing rectangles

% put your definitions there:
\startlocaldefs
\usepackage{algorithm}
\usepackage{algorithmic,amssymb}
\usepackage{graphicx,xcolor}
\usepackage{subfigure}
%\usepackage{placeins}
%\usepackage[boxed]{algorithm2e}

@@

% "Title of the paper"
\title{A significance test for forward stepwise model selection}

%%%
%%% \runtitle{} - I don't know what this is
%%%

% indicate corresponding author with \corref{}
\author{\fnms{Jonathan} \snm{Taylor}\corref{Jonathan E. Taylor}\ead[label=e1]{jonathan.taylor@stanford.edu}\thanksref{t1}
\and \fnms{Joshua} \snm{Loftus}}
\thankstext{t1}{Supported in part by NSF grant DMS 1208857 and AFOSR grant 113039.}

@@

%%%
%%% \runauthor{Taylor} - I don't know what this is
%%%

@@

%%%
%%% - Gotta change these??????????
%%%
\begin{keyword}[class=AMS]
\kwd[Primary ]{62M40}
%\kwd{}
\kwd[; secondary ]{62H35}
\end{keyword}

@@

In Table~\ref{tab:example} we see that forward stepwise chooses the
truly nonzero variables first and both $p$-values for these are small.
However, once forward stepwise begins adding noise variables, the
$\chi^2$ $p$-value remains small, potentially leading to 
incorrect inferences. The $T\chi$ and MC-estimated exact $p$-values
do not suffer from this selection effect.
%However, the MC-estimate takes substantially more computational time to evaluate than the $T\chi$ statistic.

@@

We allow forward stepwise selection to add groups of variables in each
step, not only in the case of dummy variable encoding for categorical
variables but also for any grouping purpose. For example, groups of
variables may be pre-designated factors such as expression
measurements for all genes in a single functional pathway.
To emphasize this we use $g, h$ as covariate indices rather than the usual
$i, j$ throughout. Since single variables can be considered groups of
size 1, this includes non-grouped situations as a special
case.
%Our variant of forward stepwise uses a different method than
%usual for choosing which group to add because we are using the
%same objective in the forward stepwise procedure that was used to
%derive the test statistics in penalized regression settings.

Let $y \in \real^n$ be $n$ i.i.d. measurements of the outcome variable.
Let an integer $G \geq 2$ be the number of groups of explanatory variables.
For each $1 \leq g \leq G$ the design matrix encoding the
$g$th group is the $n \times p_g$ matrix denoted $X_g$, where $p_g$ is
the number of individual variables or columns in group $g$.
% Do we need to assume $p_g \ll n$ or $\text{rank}(X_g) = p_g$?
When a group encodes a categorical variable as indicators for the
levels of that variable, by default we use the full encoding with a
column for every level. Although this introduces collinearity, our method
does not require the design matrix to have full rank.

@@

First, the user must specify the maximum number of steps allowed,
which we denote \textit{steps}.
%(consistent with the \textit{R} function for stepwise).
To enable our test statistic computations,
\textit{steps} should be at most $\min (n, G) - 1$, but it is
computationally desirable to set it as low as possible while
safely larger than the sparsity of $\beta$. 
Then forward stepwise may recover all the nonzero
coefficients of $\beta$ and terminate without performing much
additional computation.
Of course the sparsity is usually unknown, so this requires guesswork.
In our implementation we treat the active set $A$ as an ordered list to easily track the order of groups entering the model.

@@

Assuming groups of size one and $\Sigma = \sigma^2 I$, the covariance test is given at the first step by
\begin{equation}
\label{eq:covtest}
T_1 = \frac{\lambda_1(\lambda_1-\lambda_2)}{\sigma^2} \overset{n,p \to \infty}{\to} \text{Exp}(1).
\end{equation}
This is a hypothesis test for including the first variable in the LASSO
path, with large values of the test statistic being evidence against the
global null.
In \cite{tests:adaptive} it is pointed out 
\begin{equation}
\label{test:exact}
\exp\left(- \frac{\lambda_1(\lambda_1-\lambda_2)}{\sigma^2} \right) \approx \frac{1 - \Phi(\lambda_1/\sigma)}{1 - \Phi(\lambda_2 / \sigma)} \overset{D}{=} \text{Unif}([0,1]).
\end{equation}
The right hand side has an exact, finite sample null distribution.
Further, this new statistic can be understood as the survival function of $\lambda_1 = \|X^Ty\|_{\infty}$
{\em conditional} on which variable achieves $\lambda_1$ as well as its sign.
The limiting results about later steps in \cite{significance:lasso} can
%(formally)
be interpreted as
recursively applying $T_1$ to the variables that had not previously been selected by LARS. In forward stepwise with groups of size 1, we recursively apply the right hand side test statistic in the same manner. With larger groups, we use an analogous test statistic to the right hand side of \eqref{test:exact}. The test statistic is presented as an example in \cite{tests:adaptive} though we re-derive it here in simpler form. The test draws inspiration
from the group LASSO.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%% Related works %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \begin{figure}
% \label{fig:polytope}
% \begin{center}
% \resizebox{!}{3in}{\input{../figs/polytope.pdf_tex}}
% \end{center}
% A polytope selection event. Solving the LASSO \eqref{eq:lasso} for a fixed value
% of $\lambda$ yields both an active set $\hat{E}(y)$ and a collection of signs of 
% active variables $z_{\hat{E}(y)}$. The event $\{y: (\hat{E}(y), z_{\hat{E}(y)}) = (E, z_E)\}$
% can be written as a set of affine constraints.
% \end{figure}

@@

We now give the full derivation of the $p$-value, beginning with a previous work and then extending the framework there to the setting with groups.
The approach to post-selection inference in \cite{lasso:fixed}
describes the selection event for the LASSO as a convex polytope
satisfying a list of affine constraints.
%stylized in Figure \ref{fig:polytope}.
If all groups are of size 1, then
the event that we observe a given set of variables
in the forward stepwise path (with their signs as they enter) would similarly
be given by a set of affine inequalities.
%This is described in more detail in \cite{exact:lars}.
After selection, exact inference for a (randomly) chosen contrast $\eta(y)^T\mu$ of the mean vector
$\mu$ could then be accomplished by analyzing the conditional distribution
\begin{equation}
y \sim N(\mu, \sigma^2 I) \bigl| Ay \leq b.
\end{equation}
See \cite{lasso:fixed,exact:lars} for further details on this approach.

@@

%However, we must first modify our implementation of forward stepwise to prevent the possibility of including main effect groups after the main effects are already included in the model as part of an interaction group. To do this we compute the residual using the whole model in each step, rather than by subtracting the effect of the current group being included. For a set of group indices $A \subseteq \{ 1, \ldots G \}$ we write $X_A$ for the submatrix of the design matrix containing all the groups in $A$.
% \begin{algorithm}
% \DontPrintSemicolon
% \caption{\small \it Modification to compute residual from full model}
% \BlankLine
% \CommentSty{\# Everything the same as before, additionally:}\;
% %  $P_A \gets I - X_AX_A^\dagger$\;
% %  $r_s \gets P_A r_{s-1}$\;
% \If{$g^*$ is an interaction group}{
%   $g_1, g_2 \gets$ main effects groups of $g^*$\;
%   A^c \gets A^c \slash \{ g_1, g_2 \}\;
% }%\;
% %$r_s \gets y - \text{lsfit}(X_A, y)$\;
% \BlankLine
% \label{algo:fs:full}
% \end{algorithm}
% \begin{algorithm}
% \DontPrintSemicolon
% \KwData{An $n$ vector $y$ and $n \times p$ matrix $X$ of $G$ groups of variables}
% \KwResult{Active set $A$ of variable groups included in the model}
% \SetKwFunction{lsfitResidual}{lsfitResidual}
% \caption{\small \it Forward stepwise for glinternet}
% \BlankLine
% Compute design matrix~(\ref{eq:glintmat}) and expanded grouping $\tilde G$\;
% $A \gets \emptyset$\;
% $A^c \gets \{ 1, \ldots, G \}$\;
% $r_0 \gets y$\;
% \For{$s \gets 1$ \KwTo $steps$}{
%   $g^* \gets \argmax_{g \in A^c} \{ \norm{X_g^T r_{s-1}}_2 / w_g \}$\;
%   $P \gets I - X_{g^*}X_{g^*}^\dagger$\;
%   $A \gets A \cup \{ g^* \}$\;
%   $A^c \gets A^c \backslash \{ g^* \}$\;
%   \For{$g \in A^c$}{
%     $X_g \gets P X_g$\;
%   }
%   $r_s \gets P r_{s-1}$\;
% }
% \KwRet{$A$}
% \label{algo:fsgl}
% \end{algorithm}

To demonstrate this method by simulation, we constructed signals which have
the first $k/3$ main effects nonzero but with no interactions, and the
remaining $2k/3$ nonzero main effects are matched to each other to form
interactions.
%This way each nonzero main effect with any nonzero interactions has
%exactly one nonzero interaction. 
We also inflate each nonzero interaction
coefficient to be larger than the corresponding main
effect coefficients. This special case is favorable for our algorithm,
but our purpose here is merely to demonstrate the flexibility of the
hypothesis test and not to propose an optimal procedure for models with
interactions.

@@

\begin{figure}[!htp]
\begin{center}
\includegraphics[width=0.95\textwidth]{../figs/HIV_PI.pdf}
\caption{\small \it Forward stepwise results from PI dataset}
\label{fig:HIVPI}
\end{center}
\end{figure}

\begin{figure}[!htp]
\begin{center}
\includegraphics[width=0.95\textwidth]{../figs/HIV_NRTI.pdf}
\caption{\small \it Forward stepwise results from NRTI dataset}
\label{fig:HIVNRTI}
\end{center}
\end{figure}

@@

%\begin{figure}
%\begin{center}
\input{../tables/HIV_PI.tex}
%\label{tab:HIVPI}
%\end{center}
%\end{figure}

%\begin{figure}
%\begin{center}
\input{../tables/HIV_NRTI.tex}
%\label{tab:HIVNRTI}
%\end{center}
%\end{figure}

@@

\begin{figure}[!htp]
\begin{center}
\includegraphics[width=0.95\textwidth]{../figs/HIV_PI_glint.pdf}
\caption{\small \it \textsc{Glinternet} results from PI dataset}
\label{fig:HIVPI:glint}
\end{center}
\end{figure}

\begin{figure}[!htp]
\begin{center}
\includegraphics[width=0.95\textwidth]{../figs/HIV_NRTI_glint.pdf}
\caption{\small \it \textsc{Glinternet} results from NRTI dataset}
\label{fig:HIVNRTI:glint}
\end{center}
\end{figure}

@@

In this section we describe an alternative derivation following the
discussion of group LASSO in \cite{tests:adaptive}. The implementation
described here was used in all simulations.
For the rest of this section let $g = \gstar$ be the index of
the group attaining the maximum on line 3 of Algorithm~\ref{algo:fs}.
%$\lambda = \norm{X_g^Ty}_2/w_g$ be the value of the maximum, and
%$r = \text{rank}(X_g)$ be the rank of the maximizing group.
For a vector $u \in \real^p$, let $u_h$ denote the coordinates
of $u$ corresponding to the columns of group $h$ in the design matrix $X$.
We can rearrange the columns of $X$ to group these adjacently, so that
\[
u^TX^T = \begin{pmatrix} u_1^T X_1^T & u_2^T X_2^T & \cdots & u_G^TX_G^T \end{pmatrix}
\]
One step of the calculation will be to find an orthonormal basis for the
linear space $\vecsp_g = \{ u \in \real^p : u_g^T X_g^T y = 0, u_h = 0
\text{ for all } h \neq g \}$ so that we can project orthogonally to
this space. If $X_g$ is a single column and $X_g^Ty \neq 0$ (which
should be the case since $g$ maximizes the absolute value of this
quantity), then the
space is trivial and the desired orthogonal projection is the identity.
Otherwise, if $X_g$ has $p_g > 1$ columns the space $\vecsp_g$
generally has dimension $p_g-1$.

%We proceed by forming the $n \times p_g$ matrix
%$X_g - X_g X_g^T y y^T X_g/\norm{X_g^Ty}_2^2$ and performing Gram-Schmidt
%on the first $r-1$ columns. For precise theorems showing that this will
%generally work, see \cite{outerp:indep}.
We compute an orthonormal basis by Gram-Schmidt and form a
$p_g \times p_g$ matrix, which we denote $V_g$, by appending
0's in the coordinates corresponding to all groups $h \neq g$ and
an additional column of zeroes (since Gram-Schmidt only produces $p_g-1$).
We can now define the projection

@@

% \todo Clean this section \\
% \todo Make $M=\lambda_2$ a special case corollary \\

% For the sake of completeness this appendix provides full derivations
% of closed forms for the quantities required to compute our
% statistic. We require two facts from \cite{tests:adaptive}

% A point $\eta \in \K$ maximizes $f_\eta$ over a convex
% set $\K$ if and only if the following conditions hold: 
% \begin{equation}
% \label{eq:maxcond}
% \grad f_{|T_{\eta}\K} = 0, \quad
% \tf^{\eta}_{\eta} \geq \V^+_{\eta}, \quad
% \tf^{\eta}_{\eta} \leq \V^-_{\eta}, \quad \text{and} \quad
% \V^0_{\eta} \leq 0.
% \end{equation}
% The same equivalence holds true even when $\K$ is only locally
% convex.

% Write $M^{\pm}_{\eta, h}$ and $M^0_{\eta, h}$ as the corresponding suprema and infima over the restricted parameter set $S_h$. Note that the characterization of the global maximizer (Lemma 1, general paper) implies $M^0_{\eta, h} \leq 0$ and $M^+_{\eta, h} \leq M^-_{\eta, h}$ for all $h \neq g$. This will be used below to eliminate some degenerate cases of the optimization sub-problem on each $S_h$.

% Now fix $h$, let $P_g = X_gX_g^T$ and define
% \begin{align*}
% a_h &= w_h^{-1} X_h^T (I-P_g) y \\
% b_h &= w_g w_h^{-1} X_h^T P_g y / \norm{X_g^T y} \\
% c_h &= a_h^T b_h / (\norm{a_h} \norm{b_h}) \\
% K_h   &= \{ x : \norm{x} = 1, b_h^Tx > w \}
% \end{align*}
% Note that $K_h = \emptyset$ if and only if $\norm{b_h} < w$. We consider two cases. First, suppose $\norm{b_h} > w$. In this case we must rule out one sub-case, when $a_h$ is not in the polar cone of $K_h$. If this occurs, then on at least one side of the boundary of $K_h$ we have $a_h^Tx > 0$ there, so that the infimum inside $K$ is $-\infty$ but the supremum outside $K$ is $+\infty$. This violates the characterization of the global maximizer of the original process as described above, so this case cannot occur.

% Now, if $a_h$ is in the polar cone of $K_h$ (and ruling out the probability zero case where $a_h^Tx = 0$ on the boundary of $K_h$), then the numerator $a_h^Tx < 0$ on $K_h$ so there is a finite positive infimum in the interior $K_h$. Similarly there is a finite positive supremum on the interior of $K_h^c$ (the numerator is negative near the boundary of $K_h$ by continuity).

% To attain the infimum on $K_h$ requires making $x$ close to $b_h$ and simultaneously making the numerator closer to zero, so $x$ should be on the side of $b_h$ that is closer to $a_h$. To attain the supremum on $K_h^c$ requires making $x$ simultaneously close to $a_h$ (so that $a_h^Tx > 0$) and $b_h$ (to make the denominator small). In summary, both the infimum and supremum are attained with $x$ between $a_h$ and $b_h$, when the angle $\theta$ between $a_h$ and $b_h$ is larger than both the angle $\psi$ between $a_h$ and $x$ and the angle $\phi$ between $x$ and $b_h$. This allows the simplification $\phi = \theta - \psi$. This is easier to verify for the case when $\norm{b_h} < w$.

% We have verified that solving both cases requires finding the maxima of the trigonometric form $\norm{a_h} \cos (\psi) / (w - \norm{b_h} \cos(\theta - \psi))$ of the linear fraction on the interiors of $K_h$ (if it is nonempty) and $K_h^c$. The derivative is proportional to
% \[
% \frac{\norm{b_h} \sin(\theta) - w \sin(\psi)}{w - \norm{b_h} \cos(\theta - \psi)}
% \]
% Since we have ruled out the boundary of $K_h$, we can ignore the denominator and focus on critical points corresponding to the angles $\psi^{\pm}$ (symmetric about $\pi/2$) where $\sin(\psi^{\pm}) = (\norm{b_h} / w) \sin (\theta)$. That is,
% \[
% \psi^{\pm} \in \arcsin \frac{\norm{b_h}}{w} \sqrt{1-c_h^2}
% \]
% Note that these critical points only exist if $\sin(\theta) < w/\norm{b_h}$, and since we have argued that the infimum and supremum are attained it follows that this condition is met in our case. Let $\psi^+$ denote the smaller of the two angles. If $K_h$ is nonempty, then $\psi^-$ gives the infimum in $K_h$ and $\psi^+$ the supremum on $K_h^c$. This is necessary since if $x$ is in $K_h$ then its angle from $a_h$ is larger than $\pi/2$ because $a_h$ is in the polar cone of $K_h$. If $K_h$ is empty, then the supremum is still attained at $\psi^+$ since the numerator is negative at $\psi^- > \pi/2$ and the denominator is always positive.

% Finally we calculate the values of the linear fraction at $\psi^{\pm}$. Let us first record some facts:
% \begin{align*}
% 0 < \psi^+ < \pi/2, \quad \psi^- = \pi - \psi^+ \\
% \cos(\theta) = c_h, \quad \sin(\theta) = \sqrt{1-c_h^2} \\
% \sin(\psi^+) = \sin(\psi^-) = \frac{\norm{b_h}}{w} \sin(\theta) \\
% \cos(\psi^+) = -\cos(\psi^-) = \sqrt{1 - (\norm{b_h}/w)^2(1-c_h^2)}
% \end{align*}
% Using the angle difference formula,
% \begin{align*}
% \cos(\theta - \psi^+) & = \frac{\norm{b_h}}{w} (1-c_h^2) +  c_h \cos(\psi^+) \\
% & = \frac{\norm{b_h}}{w} (1-c_h^2) +  c_h \cos(\psi^+)
% \end{align*}
% Hence
% \begin{align*}
% M^+_{\eta, h} & = \frac{ \norm{a_h} c_h }{ w - (\norm{b_h}^2/w) (1-c_h^2) - \norm{b_h} c_h \cos(\psi^+) } \\
% % & = 
% % & = \frac{ w \norm{a} a_h^Tb_h / \norm{b_h} }{ w^2 - \norm{a}\norm{b}^2 + (a^Tb)/\norm{a} - a^Tb \cos \psi^+ }
% \end{align*}
% And $M^-_{\eta, h}$ is $\infty$ when $\norm{b_h} < w$ and otherwise given by
% \begin{align*}
% M^+_{\eta, h} & = \frac{ \norm{a_h} c_h }{ w - (\norm{b_h}^2/w) (1-c_h^2) + \norm{b_h} c_h \cos(\psi^+) } \\
% \end{align*}

% I haven't been able to algebraically simplify these yet to a form that allows my $\lambda_2$ proof below to go through. The version below just assumes $\norm{b_h} < w = 1$.

% \hrule
% \hrule
% We can rewrite this by rationalizing the denominator:
% \[
% M =
% \frac
% { a_h^T b_h +  \sqrt{ a_h^T a_h (1-b_h^Tb_h) + (a_h^T b_h)^2 }  }
% { 1-b_h^Tb_h }
% \]
% Leaving $M$ in this form we now consider $\lambda_2$. The KKT conditions for the group LASSO problem give (directly from Yuan and Lin)
% \begin{align*}
% \norm{ X_g^T(y - X \beta) } & \leq \lambda w_g \quad \forall \beta_g \neq 0 \\
% \beta_g & = \left( 1 - \frac{ \lambda w_g }{ \norm{S_g} } \right)_+ S_g
% \end{align*}
% where $S_g = X_g^T(y - X \beta_{-g})$. These expressions simplify for $\lambda_2 \leq \lambda < \lambda_1$. Letting $g$ be the index of the first group to enter we now have $S_g = X_g^Ty$ 
% \begin{align*}
% \beta_g & = \left( 1 - \frac{ \lambda w_g }{ \norm{X_g^Ty} } \right) X_g^Ty \\
% \lambda w_h & \geq \norm{ X_h^Ty - X_h^T P_g y (1-\lambda w_g / \norm{X_g^Ty})  } \quad \forall h \neq g
% \end{align*}
% Now let $\lambda = \lambda_2$, so the inequality above is strict for all other groups and equality is attained for the second group to enter. Hence $\lambda_2$ satisfies
% \begin{align*}
% \lambda_2 & = \max_{h \neq g} 
% \norm{ X_h^T(I - P_g) y - \lambda_2 ( w_g X_h^T P_g y / \norm{X_g^Ty} )  } / w_h\\
% & = \max_{h \neq g} \norm{ a_h - \lambda_2 b_h } 
% \end{align*}

% Write $\lambda_{2,h}$ as the positive root of the quadratic equation obtained by squaring the norm, so $\lambda_{2,h}^2 = \norm{ a_h - \lambda_2 b_h }^2$. Then $\lambda_2 = \max_{h \neq g} \lambda_{2,h}$. Solving this (note that $b_h^Tb_h < 1$), we find
% \[
% \lambda_{2,h} = \frac
% { a_h^T b_h +  \sqrt{ a_h^T a_h (1-b_h^Tb_h) + (a_h^T b_h)^2 }  }
% { 1-b_h^Tb_h }
% \]
% Hence when $h$ is the index of the second group to enter, we have $\lambda_2 = \lambda_{2,h} = M$.

>>>>

% put your definitions there:
\startlocaldefs
\usepackage{algorithm}
\usepackage{algorithmic,amssymb}
\usepackage{graphicx,xcolor}
\usepackage{subfigure}

@@

\title{A significance test for forward stepwise model selection}

%%% \runtitle{} - I don't know what this is

% indicate corresponding author with \corref{}
\author{\fnms{Joshua} \snm{Loftus}
\and \fnms{Jonathan} \snm{Taylor}\corref{Jonathan E. Taylor}\ead[label=e1]{jonathan.taylor@stanford.edu}\thanksref{t1}}
\thankstext{t1}{Supported in part by NSF grant DMS 1208857 and AFOSR grant 113039.}

@@

%%% \runauthor{Taylor} - I don't know what this is

@@

%%% - Check these
\begin{keyword}[class=AMS]
\kwd[Primary ]{62M40}
%\kwd{}
\kwd[; secondary ]{62H35}
\end{keyword}

@@

In Table~\ref{tab:example} we see that forward stepwise chooses the
truly nonzero variables first and both $p$-values for these are small.
However, once forward stepwise begins adding noise variables, the
$\chi^2$ $p$-value remains small, potentially leading to 
incorrect inferences. The $T\chi$ and MC-estimated exact $p$-values
do not suffer from this selection effect.
However, the MC-estimate takes substantially more computational time to evaluate than the $T\chi$ statistic.

@@

We allow forward stepwise selection to add groups of variables in each
step, not only in the case of dummy variable encoding for categorical
variables but also for any grouping purpose. For example, groups of
variables may be pre-designated factors such as expression
measurements for all genes in a single functional pathway.
To emphasize this we use $g, h$ as covariate indices rather than the usual
$i, j$ throughout. Since single variables can be considered groups of
size 1, this includes non-grouped situations as a special
case.

Let $y \in \real^n$ be $n$ i.i.d. measurements of the outcome variable.
Let an integer $G \geq 2$ be the number of groups of explanatory variables.
For each $1 \leq g \leq G$ the design matrix encoding the
$g$th group is the $n \times p_g$ matrix denoted $X_g$, where $p_g$ is
the number of individual variables or columns in group $g$.
When a group encodes a categorical variable as indicators for the
levels of that variable, by default we use the full encoding with a
column for every level. Although this introduces collinearity, our method
does not require the design matrix to have full rank.

@@

First, the user must specify the maximum number of steps allowed,
which we denote \textit{steps}.
To enable our test statistic computations,
\textit{steps} should be at most $\min (n, G) - 1$, but it is
computationally desirable to set it as low as possible while
safely larger than the sparsity of $\beta$. 
Then forward stepwise may recover all the nonzero
coefficients of $\beta$ and terminate without performing much
additional computation.
Of course the sparsity is usually unknown, so this requires guesswork.
In our implementation we treat the active set $A$ as an ordered list to easily track the order of groups entering the model.

@@

Assuming groups of size one and $\Sigma = \sigma^2 I$, the covariance test is given at the first step by
\begin{equation}
\label{eq:covtest}
T_1 = \frac{\lambda_1(\lambda_1-\lambda_2)}{\sigma^2} \overset{n,p \to \infty}{\to} \text{Exp}(1).
\end{equation}
This is a hypothesis test for including the first variable in the LASSO
path, with large values of the test statistic being evidence against the
global null.
In \cite{tests:adaptive} it is pointed out 
\begin{equation}
\label{test:exact}
\exp\left(- \frac{\lambda_1(\lambda_1-\lambda_2)}{\sigma^2} \right) \approx \frac{1 - \Phi(\lambda_1/\sigma)}{1 - \Phi(\lambda_2 / \sigma)} \overset{D}{=} \text{Unif}([0,1]).
\end{equation}
The right hand side has an exact, finite sample null distribution.
Further, this new statistic can be understood as the survival function of $\lambda_1 = \|X^Ty\|_{\infty}$
{\em conditional} on which variable achieves $\lambda_1$ as well as its sign.
The limiting results about later steps in \cite{significance:lasso} can
be interpreted as
recursively applying $T_1$ to the variables that had not previously been selected by LARS. In forward stepwise with groups of size 1, we recursively apply the right hand side test statistic in the same manner. With larger groups, we use an analogous test statistic to the right hand side of \eqref{test:exact}. The test statistic is presented as an example in \cite{tests:adaptive} though we re-derive it here in simpler form. The test draws inspiration
from the group LASSO.

@@

@@

We now give the full derivation of the $p$-value, beginning with a previous work and then extending the framework there to the setting with groups.
The approach to post-selection inference in \cite{lasso:fixed}
describes the selection event for the LASSO as a convex polytope
satisfying a list of affine constraints.
If all groups are of size 1, then
the event that we observe a given set of variables
in the forward stepwise path (with their signs as they enter) would similarly
be given by a set of affine inequalities.
After selection, exact inference for a (randomly) chosen contrast $\eta(y)^T\mu$ of the mean vector
$\mu$ could then be accomplished by analyzing the conditional distribution
\begin{equation}
y \sim N(\mu, \sigma^2 I) \bigl| Ay \leq b.
\end{equation}
See \cite{lasso:fixed,exact:lars} for further details on this approach.

@@

To demonstrate this method by simulation, we constructed signals which have
the first $k/3$ main effects nonzero but with no interactions, and the
remaining $2k/3$ nonzero main effects are matched to each other to form
interactions.
We also inflate each nonzero interaction
coefficient to be larger than the corresponding main
effect coefficients. This special case is favorable for our algorithm,
but our purpose here is merely to demonstrate the flexibility of the
hypothesis test and not to propose an optimal procedure for models with
interactions.

@@

\begin{figure}[!htp]
\begin{center}
\includegraphics[width=0.95\textwidth]{../figs/HIV_PI_plot.pdf}
\caption{\small \it Forward stepwise results from PI dataset}
\label{fig:HIVPI}
\end{center}
\end{figure}

\begin{figure}[!htp]
\begin{center}
\includegraphics[width=0.95\textwidth]{../figs/HIV_NRTI_plot.pdf}
\caption{\small \it Forward stepwise results from NRTI dataset}
\label{fig:HIVNRTI}
\end{center}
\end{figure}

@@

\input{../tables/HIV_PI.tex}

\input{../tables/HIV_NRTI.tex}

@@

\begin{figure}[!htp]
\begin{center}
\includegraphics[width=0.95\textwidth]{../figs/HIV_PI_glint_plot.pdf}
\caption{\small \it \textsc{Glinternet} results from PI dataset}
\label{fig:HIVPI:glint}
\end{center}
\end{figure}

\begin{figure}[!htp]
\begin{center}
\includegraphics[width=0.95\textwidth]{../figs/HIV_NRTI_glint_plot.pdf}
\caption{\small \it \textsc{Glinternet} results from NRTI dataset}
\label{fig:HIVNRTI:glint}
\end{center}
\end{figure}

@@

In this section we describe an alternative derivation following the
discussion of group LASSO in \cite{tests:adaptive}. The implementation
described here was used in all simulations.
For the rest of this section let $g = \gstar$ be the index of
the group attaining the maximum on line 3 of Algorithm~\ref{algo:fs}.
For a vector $u \in \real^p$, let $u_h$ denote the coordinates
of $u$ corresponding to the columns of group $h$ in the design matrix $X$.
We can rearrange the columns of $X$ to group these adjacently, so that
\[
u^TX^T = \begin{pmatrix} u_1^T X_1^T & u_2^T X_2^T & \cdots & u_G^TX_G^T \end{pmatrix}
\]
One step of the calculation will be to find an orthonormal basis for the
linear space $\vecsp_g = \{ u \in \real^p : u_g^T X_g^T y = 0, u_h = 0
\text{ for all } h \neq g \}$ so that we can project orthogonally to
this space. If $X_g$ is a single column and $X_g^Ty \neq 0$ (which
should be the case since $g$ maximizes the absolute value of this
quantity), then the
space is trivial and the desired orthogonal projection is the identity.
Otherwise, if $X_g$ has $p_g > 1$ columns the space $\vecsp_g$
generally has dimension $p_g-1$.

We compute an orthonormal basis by Gram-Schmidt and form a
$p_g \times p_g$ matrix, which we denote $V_g$, by appending
0's in the coordinates corresponding to all groups $h \neq g$ and
an additional column of zeroes (since Gram-Schmidt only produces $p_g-1$).
We can now define the projection

@@
